Sketch Recognition and Understanding

### 2009-09-06 draft/System.tex
after the changes in examples

SVM have been applied ot the a largest MNIST a Handwritten digit recognition dataset of 60000 training set and 10000 test set\cite{IjdarSherifPaper}. Sherif and Ezzat has performed a comparison between SVM and several other techniques on the MNIST dataset. They also compared the same classification methods on arabic handwritten digits (MADBASE) \cite{ADBase9,IjdarSherifPaper}. Their results shows that SVM classifier outperform all other techniques on both MNIST and MADBase datasets and achieved recognition rate of over 98\%. Similar results was reported by \cite{empiricalcomp11,SVMInvariantComDecoste02} on the MNIST dataset. Ther are many other successful application SVM. They have proven to be robust to noise and perform well on many tasks \cite{empiricalcomp11,Scholkopf97comparingsupport}.

This superior performance because they may eliminates many of problem experienced with other inference methodoliges with other inference methodologies like neural network and decision trees. The first problem that is eliminated was the problems with local minima. SVM can construct highly nonlinear classification and regression function without working about getting stuck at local minima. Another problem SVM eliminated is the number of parameter that are picked ot generate the model. For example, SVM classifier with RBF  kernel will only need 2 parameters to set the model the $C$ cost parameter and $\sigma$ parameter of the RBF kernel. The last reason is the final result are sable, reproducible and largely independent of the specific algorithm used to  optimize the SVM model. If two users apply the same SVM model with the same parameters to the same data, they will et the same solution module numeric issues. Compare this with neural networks where the results are dependent on the particular algorithms and starting point used.

@@

\subsubsection {SVM Implementations}
To solve SVM approach require solving a QP or LP problem.  Mathematical programming have extensively studied the LP and QP type problems. These techniques can be divided into three categories, methods in which evaluate and discard kernel components during learning. Decomposition methods in which an evolving subset of data are used and new optimization approaches that specifically utilize the structure of the SVM problem.

@@

The third approach is to directly attack the SVM  problem from and optimization perspective and create algorithms that explicitly exploits the structure of teh problem. Keerthi et al. \cite{fastSVM6} proposed a very effective algorithm based on the dual geometry of finding the two closest points in the convex hulls.

@@

% whate is support vector machines 
%%The classifcation problem can be restricted to consideration of the two-class problem
%without loss of generality. In this problem the goal is to separate the two classes by a
%function which is induced from available examples. The goal is to produce a classifier
%that will work well on unseen examples, i.e. it generalises well. Consider the example
%in Figure 2.1. Here there are many possible linear classifiers that can separate the data,
%but there is only one that maximises the margin (maximises the distance between it
%and the nearest data point of each class). This linear classifier is termed the optimal
%separating hyperplane. Intuitively, we would expect this boundary to generalise well as
%opposed to the other possible boundaries.
%
%Let $x_i$ , $i$ = $1.2 \dots,N$ , be the feature vectors of the training set, $X$ . These
%belong to either of two classes that are linearly spereable $\omega_1=1$ and  $\omega_2=-1$.
%The goal, is to design a hyperplane which solve the equation 
%\begin{equation}
%g(x) = \omega^{T} x + \omega g = 0
%\label{eq:svmhyper}
%\end{equation}
 % Vladimir N. Vapnik. Statistical Learning Theory. John Wiley & Sons, Inc., New
% York, NY, September 1998.

@@

\section{Step by Step Example}
\label{secstepExample}
In this section a step by step example is presented to clarify the presented system. Let us make an assumption that the user will draw a clock using two strokes. Figure \ref{fig:clock} show the two stroke and how the user draw them. The next steps demonstrate how output of the system after each step. The first section \ref{sec:stepseg} describe how the segmentation is done step by step. Section \ref{sec:steprec} shows the details of feature calculation and the classification process.

@@

\begin{figure}
	\centering
	\subfigure[First Stroke]{\label{fig:clockstroke1}
		\includegraphics[scale=0.7]{afterDefense/clockstroke1.jpg.eps} }

		\subfigure[Second Stroke]{	\label{fig:clockstroke2}
		\includegraphics[scale=0.7]{afterDefense/ClockComplete.jpg.eps}}

		\subfigure[Completed symbol]{
		\label{fig:Clock1}\includegraphics[scale=0.7]{afterDefense/Clock1.jpg.eps}	}
	\caption{Stroke Drawing Sequence} a) The user first draws the ellipse b) The square wave is drawn inside the ellipse c) The arrows in the symbols illustrate how the user has drawn the symbol
	\label{fig:clock}
\end{figure}

In this step the user draw one or more stroke to draw a symbol. Figure \ref{fig:clock} illustrate the sequence and direction of user draw. In this section, each stroke user draw will be labeled as it illustrated in Figure \ref{fig:clockstroke1} and Figure \ref{fig:clockstroke2} . Table \ref{tab:StepsStroke1} shows the detailed output of the program at each step. The input of this step is the stroke points and the ouptut is a list of segments that wil be used in the recognition process. The algorithm that compute the segmentation is demonstrated as follows:
\begin{enumerate}

	\item Extracting $P_dp$ as in Section \ref{sec:Preprocessing}
\begin{verbatim}
  1. For each point in Stroke do the follwoing:
  		1.a Compute distance from previous point.
  		1.b Compute Time difference from previous point and add it to Time_list.  
  		1.c Compute Velocity and add it to Vel_list. 
  		1.d Compute Direction and add it to Dir_list.
  	 	1.e Compute Curvature and add it to Cur_list.(Eq.4.1) 
  	end (for loop)
  2. For each  list computed in previous step do the follwing:   
  	 2.a Compute average of all points. 
  	 2.b set average as threshold.  
  	 2.c Create region from continous points larger than threshold. 
  	 2.d for each region do the following. 
  	 	   2.d.1 find the maximum 
  	 	   2.d.1 Add point to P_dp 
  	  end
     end 
\end{verbatim}	
\item Ellipse Fitting: which is an attempt to fit as an ellipse as in  Section \ref{sec:EllipseDetection} 
%he axes of the ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box where $width$ is width and $height$ is height of the bonding box.
\begin{verbatim}
1. Initialize the axes of ellipse  and center by bounding box information
2. Repeat the follwoing for few loops: 
	2.a Compute error in Equation  4.2 
	2.b Update center of ellipse and axis from error. 
3. Compute percent in Equation 4.4 
4. Computer eff in Equation 4.3 
5. if eff < 0.2  then fit as ellipse otherwise move to segment. 
\end{verbatim}

\item Segmentation ALS1 as explained in Section \ref{sec:SwarmSegmentation}. The alogrithm will be similar to the PSO algoirhtm  in section \ref{sec:ParticleSwarmAlgorithm} 
\begin{verbatim}
Initialize 
  1. Generate N particles with M dimensions (P_i...P_N) 
  2. Generate velocities for each particle for M dimension randomly. 
  3. Repeat till Maximum number of iteration OR ErrorThreshold  
         For each particle P_i do the following 
         3.a   Compute fitness  (fitness) 
         				3.a.1 Compute the Error of approximation as in Eq.4.5
         				3.a.2. Computer the fitness function as in Eq. 4.4 
         3.b   Determine Lbest  
               if (lbest_fitness<fitness)
                  3.b.1   lbest=P_i
         3.c      Determine Gbest.
               if (lGest_fitness<fitness)
                  3.c.1   lGest=P_i
         3.d   Compute the Velocity for each dimension using Eq.3.2 
         3.e   Bound Velocity using Eq.3.4
         3.f   Update particle values using Eq. 3.3
         3.g   Refine the location of the particle 
         end .
     end 
4. return Gbest
\end{verbatim}	
\item Segmentation ALS2  as explained in Section \ref{sec:PolygonDivisionAlgorithm}
\begin{verbatim}
Initialize 
  1. Generate N particles with M dimensions (P_i...P_N) 
  2. Generate velocities for each particle for M dimension randomly. 
  3. Repeat till Maximum number of iteration OR ErrorThreshold  
         For each particle P_i do the following 
         3.a   Compute fitness  (fitness) 
              3.a.1 For each Segment in the particle do the following
         				    3.a.1.I   Error of approximation as Line (Eq. 4.7 - 4.9 )
         			      3.a.1.II  Error of approximation as circular arc (Eq. 4.10 - 4.19 )
         				    3.a.1.III Computer minimum of both approximation 
         			3.a.2 Compute Error of the particle as in Eq.  4.20
         			3.a.3 Compute fitness of the particle as in Eq. 4.21
         3.b   Determine Lbest  
               if (lbest_fitness<fitness)
                  3.b.1   lbest=P_i
         3.c      Determine Gbest.
               if (lGest_fitness<fitness)
                  3.c.1   lGest=P_i
         3.d   Compute the Velocity for each dimension using Eq.3.2 
         3.e   Bound Velocity using Eq.3.4
         3.f   Update particle values using Eq. 3.3
         3.g   Refine the location of the particle 
         end .
     end 
4. return Gbest
\end{verbatim}	
\item Append best segmentation to List of segmentation ($L_s$) of this symbols. 
\end{enumerate}

\begin{landscape}
\begin{scriptsize}
%\begin{table}
	%\centering 
%\scalebox{0.6}{		
\begin{longtable}{|p{2cm}|p{2cm}|p{2cm}|p{2cm}|p{13cm}|}
\caption{Detailed Output of System in Each Step of Segmentation}
\label{tab:StepsStroke1} \\

\hline 
\multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
\multicolumn{1}{p{2cm}|}{\textbf{Description}} &
\multicolumn{1}{p{2cm}|}{\textbf{Inputs}} &
\multicolumn{1}{p{2cm}|}{\textbf{Details}} &
\multicolumn{1}{p{13cm}|}{\textbf{Output}} 
\\ \hline 
\endfirsthead

\hline
\multicolumn{3}{c}%
{{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\
\multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
\multicolumn{1}{p{2cm}|}{\textbf{Description}} &
 \multicolumn{1}{p{2cm}|}{\textbf{Inputs}} &
 \multicolumn{1}{p{2cm}|}{\textbf{Details}} &
 \multicolumn{1}{p{13cm}|}{\textbf{Output}} 
\\ \hline 
\endhead

%Step & Discription of Step & Input & Steps & Output \\ \hline
Stroke 1 & Extracting point and computing boundary data & -- &  see Section \ref{sec:Preprocessing} & 
\begin{scriptsize}
Stroke has 62 points = [P(198.0,153.0), P(197.0,150.0), P(197.0,145.0), P(198.0,140.0), P(203.0,125.0), P(205.0,120.0) , P(205.0,117.0), P(210.0,111.0) , P(214.0,103.0) , P(219.0,97.0), P(225.0,89.0) , P(229.0,85.0) , P(242.0,77.0) , P(247.0,73.0) , P(275.0,64.0) , P(287.0,62.0) , P(325.0,62.0) , P(333.0,63.0) , P(362.0,68.0) , P(372.0,72.0) , P(391.0,77.0) , P(406.0,83.0) , P(424.0,91.0) , P(440.0,100.0), P(474.0,133.0) , P(485.0,149.0) , P(492.0,164.0) , P(499.0,182.0) , P(501.0,191.0) , P(501.0,207.0) , P(499.0,222.0) , P(496.0,239.0) , P(471.0,292.0) , P(467.0,302.0) , P(456.0,315.0) , P(441.0,329.0) , P(427.0,339.0) , P(418.0,344.0) , P(398.0,354.0) , P(336.0,371.0) , P(283.0,376.0) , P(263.0,378.0) , P(242.0,378.0) , P(232.0,376.0) , P(213.0,373.0) , P(194.0,367.0) , P(177.0,362.0) , P(138.0,343.0) , P(129.0,336.0) , P(124.0,331.0) , P(120.0,321.0) , P(115.0,310.0) , P(113.0,297.0) , P(112.0,284.0) , P(114.0,256.0) , P(121.0,219.0) , P(128.0,202.0) , P(137.0,184.0) , P(143.0,172.0) , P(154.0,161.0) , P(172.0,146.0) , P(172.0,146.0)]

 Bonding box is  = Corner P(112.0, 62.0), w= 389.0, h=  316.0
 number of point in this stroke is   62
\end{scriptsize}
 \\ \hline
Possible Dominate point Extraction & Computing and for details see Section\ref{sec:Preprocessing} &  Stroke points  &   &    Number of $P_{dp}$  = 11
 Pdp = [ 2 (197.0 , 145.0 ), 4 (203.0 , 125.0 ), 3 (198.0 , 140.0 ), 22 (424.0 , 91.0 ), 19 (372.0 , 72.0 ), 26 (492.0 , 164.0 ), 28 (501.0 , 191.0 ), 9 (219.0 , 97.0 ), 31 (496.0 , 239.0 ), 0 (198.0 , 153.0 ), 41 (263.0 , 378.0 )]

\\ \hline 
Segmentation & Segmentation of Ellipse & $P_{dp}$   &  Ellipse fitting  & 
 Segmented as Ellipse with dimensions  a = 194.5 ,b = 194.5 , Center( 306.5,  220.0 )   with Error =  2.276869953758031 , Percent =0.9093857875960022
 certainty  =  0.1997008626019765 
\\ \hline

>>>>

SVM have been applied ot the a largest MNIST a Handwritten digit recognition dataset of 60000 training set and 10000 test set\cite{IjdarSherifPaper}. Sherif and Ezzat has performed a comparison between SVM and several other techniques on the MNIST dataset. They also compared the same classification methods on Arabic handwritten digits (MADBASE) \cite{ADBase9,IjdarSherifPaper}. Their results shows that SVM classifier outperform all other techniques on both MNIST and MADBase datasets and achieved recognition rate of over 98\%. Similar results was reported by \cite{empiricalcomp11,SVMInvariantComDecoste02} on the MNIST dataset. There are many other successful application SVM. They have proven to be robust to noise and perform well on many tasks \cite{empiricalcomp11,Scholkopf97comparingsupport}.

This superior performance because they may eliminates many of problem experienced with other inference methodologies with other inference methodologies like neural network and decision trees. The first problem that is eliminated was the problems with local minima. SVM can construct highly nonlinear classification and regression function without working about getting stuck at local minima. Another problem SVM eliminated is the numbers of parameter that are picked to generate the model. For example, SVM classifier with RBF kernel will only need 2 parameters to set the model the $C$ cost parameter and $\sigma$ parameter of the RBF kernel. The last reason is the final result are sable, reproducible and largely independent of the specific algorithm used to  optimize the SVM model. If two users apply the same SVM model with the same parameters to the same data, they will et the same solution module numeric issues. Compare this with neural networks where the results are dependent on the particular algorithms and starting point used.

@@

\subsubsection {SVM Implementations}
To solve SVM approach requires solving a QP or LP problem.  Mathematical programming have extensively studied the LP and QP type problems. These techniques can be divided into three categories, methods in which evaluate and discard kernel components during learning. Decomposition methods in which evolving subsets of data are used and new optimization approaches that specifically utilize the structure of the SVM problem.

@@

The third approach is to directly attack the SVM problem from and optimization perspective and create algorithms that explicitly exploits the structure of ten problem. Keerthi et al. \cite{fastSVM6} proposed a very effective algorithm based on the dual geometry of finding the two closest points in the convex hulls.

@@

% whate is support vector machines 
%%The classifcation problem can be restricted to consideration of the two-class problem
%without loss of generality. In this problem the goal is to separate the two classes by a
%function which is induced from available examples. The goal is to produce a classifier
%that will work well on unseen examples, i.e. it generalizes well. Consider the example
%in Figure 2.1. Here there are many possible linear classifiers that can separate the data,
%but there is only one that maximizes the margin (maximizes the distance between it
%and the nearest data point of each class). This linear classifier is termed the optimal
%separating hyper plane. Intuitively, we would expect this boundary to generalize well as
%opposed to the other possible boundaries.
%
%Let $x_i$ , $i$ = $1.2 \dots,N$ , be the feature vectors of the training set, $X$ . These
%belong to either of two classes that are linearly spereable $\omega_1=1$ and  $\omega_2=-1$.
%The goal, is to design a hyperplane which solve the equation 
%\begin{equation}
%g(x) = \omega^{T} x + \omega g = 0
%\label{eq:svmhyper}
%\end{equation}
 % Vladimir N. Vapnik. Statistical Learning Theory. John Wiley & Sons, Inc., New
% York, NY, September 1998.

@@

\section{Step by Step Example}
\label{secstepExample}
In this section a step by step example is presented to clarify the presented system. Let us make an assumption that the user will draw a clock using two strokes. Figure \ref{fig:clock} show the two strokes and how the user draws them. The next steps demonstrate how output of the system after each step. The first section \ref{sec:stepseg} describe how the segmentation is done step by step. Section \ref{sec:steprec} shows the details of feature calculation and the classification process. 
\begin{figure}
	\centering
	\subfigure[First Stroke]{\label{fig:clockstroke1}
		\includegraphics[scale=0.7]{afterDefense/clockstroke1.jpg.eps} }
		\subfigure[Second Stroke] {	\label{fig:clockstroke2}
		\includegraphics[scale=0.7]{afterDefense/clockstroke2.jpg.eps}}		
		\subfigure[Completed symbol]{
		\label{fig:Clock1}\includegraphics[scale=0.7]{afterDefense/ClockDirections.jpg.eps}	}
	\caption{Stroke Drawing Sequence} a) The user first draws the ellipse b) The square wave is drawn inside the ellipse c) The final symbol the user draws along with arrows to illustrate how the directions the user used while drawing the symbol. 
	\label{fig:clock}
\end{figure}

@@

In this step the user draw one or more stroke to draw a symbol. Figure \ref{fig:clock} illustrate the sequence and direction of user draw. In this section, each stroke user draws will be labeled as it illustrated in Figure \ref{fig:clockstroke1} and Figure \ref{fig:clockstroke2}.  Table \ref{tab:StepsStroke1} shows the detailed output of the program at each step. The input of this step is the a set of strokes, each stroke has a set of  points and the output is a list of segments that will be used in the recognition process. More detailed example and algorithms are demonstrated in the Appendex \ref{ChapterstepExample} \\

For each stroke the user draw the system segments the stroke using the following procedure

\begin{enumerate}
\item Extracting $P_dp$ as in Section \ref{sec:Preprocessing} and Algorithm \ref{extractpdp}.
\item Ellipse Fitting: which is an attempt to fit as an ellipse as in details in Section \ref{sec:EllipseDetection} and Algorithm \ref{ellipseFitAlg} in Appendex \ref{ChapterstepExample}.
\item Segmentation ALS1 as explained in Section \ref{sec:SwarmSegmentation}. The alogrithm will be similar to the PSO algorithm  in section \ref{sec:ParticleSwarmAlgorithm}. (see Algorithm \ref{AlgS1alg} in Appendex \ref{ChapterstepExample})
%he axes of the ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box where $width$ is width and $height$ is height of the bonding box.
\item Segmentation ALS2 as explained in Section \ref{sec:PolygonDivisionAlgorithm}. ( see Algorithm \ref{AlgS2alg} in Appendex \ref{ChapterstepExample} ). 
\item Append best segmentation to List of segmentation ($L_s$) of this symbols. 
\end{enumerate}

\begin{landscape}
The steps of the first stroke (details of the ouputs are in  Appendex \ref{ChapterstepExample} in Table \ref{tab:StepsStroke1Details}):
\begin{scriptsize}
%\begin{table}
	%\centering 
%\scalebox{0.6}{		
\begin{longtable}{|p{2cm}|p{2cm}|p{5cm}|p{2cm}|p{5cm}|}
\caption{ Output of System in Each Step of Segmentation of First Stroke}
\label{tab:StepsStroke1} \\

\hline 
\multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
\multicolumn{1}{p{2cm}|}{\textbf{Description}} &
\multicolumn{1}{p{5cm}|}{\textbf{Inputs}} &
\multicolumn{1}{p{2cm}|}{\textbf{Details}} &
\multicolumn{1}{p{5cm}|}{\textbf{Output}} 
\\ \hline 
\endfirsthead

\hline
\multicolumn{3}{c}%
{{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\
\multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
\multicolumn{1}{p{2cm}|}{\textbf{Description}} &
 \multicolumn{1}{p{5cm}|}{\textbf{Inputs}} &
 \multicolumn{1}{p{2cm}|}{\textbf{Details}} &
 \multicolumn{1}{p{5cm}|}{\textbf{Output}} 
\\ \hline 
\endhead

%Step & Discription of Step & Input & Steps & Output \\ \hline
Stroke 1 & Extracting point and computing boundary data & -- &   &

	\includegraphics[scale=0.5]{afterDefense/clockstroke1.eps}

  \\ \hline
Possible Dominate point Extraction & for Computing details see Section\ref{sec:Preprocessing} &

	\includegraphics[scale=0.5]{afterDefense/clockstroke1.eps}
  &   &    
\\ \hline 
Segmentation & Segmentation of Ellipse & $P_{dp}$   &  Ellipse fitting  & 
The circle is   a = 103.05715108084719 ,b = 103.06851874747797 , Center( 283.03365107691775,  181.05035287667172 )   with Error =  1.9471676673301757 , Percent =1.0695193086531243
eff  = 0.27463462099275116

 \includegraphics[scale=0.5]{afterDefense/ellipse.jpg.eps}
 \\ \hline
 	\end{longtable}
 	\newpage
 	The steps of the second stroke:(details of the ouputs are in  Appendex \ref{ChapterstepExample} in Table \ref{tab:StepsStroke1Details})

\begin{longtable}{|p{2cm}|p{2cm}|p{5cm}|p{2cm}|p{5cm}|}
\caption{ Output of System in Each Step of Segmentation of Second Stroke}
\label{tab:StepsStroke2} \\

\hline 
\multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
\multicolumn{1}{p{2cm}|}{\textbf{Description}} &
\multicolumn{1}{p{5cm}|}{\textbf{Inputs}} &
\multicolumn{1}{p{2cm}|}{\textbf{Details}} &
\multicolumn{1}{p{5cm}|}{\textbf{Output}} 
\\ \hline 
\endfirsthead

\multicolumn{3}{c}%
{{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\ \hline
\multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
\multicolumn{1}{p{2cm}|}{\textbf{Description}} &
 \multicolumn{1}{p{5cm}|}{\textbf{Inputs}} &
 \multicolumn{1}{p{2cm}|}{\textbf{Details}} &
 \multicolumn{1}{p{5cm}|}{\textbf{Output}} 
\\ \hline 
\endhead

Stroke 2 & Extracting point and computing boundary data & -- &  see Section \ref{sec:Preprocessing} &
 	\includegraphics[scale=0.5]{afterDefense/clockstroke2.jpg.eps} \\ \hline
  & Possible Dominate point Extraction for details see Section\ref{sec:Preprocessing} &   	\includegraphics[scale=0.5]{afterDefense/clockstroke2.jpg.eps}  & 1. compute the $P_{pd}$  & 
    \\ \hline

Segmentation & Segmentation of Ellipse & $P_{dp}$   &  Ellipse fitting  & 
  The Ellipse is   a = 125.5 ,b = 125.5 , Center( 320.5,  235.0 )   with Error =  7.074966523515405 , Percent =1.1481777694994926 certainty  =  0.08114368920921669  
 \\ \hline
 Segmentation & PSO Segmentation ALgS1 & $P_{dp}$  & ALgS1   & Global best fitness is 0.875 and error value is 729.248312158671 
 \includegraphics[scale=0.7]{afterDefense/Stroke2Seg1labeled.jpg.eps}
 \\ \hline 
  Segmentation & PSO Segmentation ALgS2 & $P_{dp}$  & ALgS2   & Error value is 8.61499750333895
 \includegraphics[scale=0.7]{afterDefense/Stroke2Segmentaion.jpg.eps}
 \\ \hline 
Choosing best fit & Compare between ALgS1 and AlgS2 &  First PSO and Second PSO  &   & The error ALgS1 780.6652975044847  Algs2 161.60692855994625    Best = ALgS2

\includegraphics[scale=0.7]{afterDefense/Stroke2Seg2labeled.jpg.eps}	
  \\ \hline 
		\end{longtable}
\end{scriptsize}
\end{landscape}
The final output of this step is list $L_s$ of segments There are  9  segments in this symbol (\ref{fig:FinalSegmentationLabeled})

\begin{itemize}
	\item    Segment S0  = Ellipse ( Stroke 1)
   \item Segment S1  =   Line ( Stroke 2)
     \item  Segment S2 = Line ( Stroke 2)
     \item  Segment S3  =  Line ( Stroke 2)
     \item Segment S4  =   Line ( Stroke 2)
  \item    Segment S5  =  Curve  ( Stroke 2)
     \item  Segment S6  =  Curve  ( Stroke 2)
     \item  Segment S7  =  Line   ( Stroke 2)
     \item  Segment S8  =  Line   ( Stroke 2)
\end{itemize}
\begin{figure}
	\centering
 		\includegraphics[scale=0.65]{afterDefense/FinalSegmentation.jpg.eps}

		\label{fig:FinalSegmentationLabeled}
	\caption{Final Segmentation of Symbol} The segments are labeled on the symbol. 
\end{figure}

\subsection{Step by Step Recognition}
\label{sec:steprec}

\begin{enumerate}
	\item Computing Structural and geometrical Features (FS1). This step computes number of parallel, perpendicular line and similar structural features [10 features]. Algorithm \ref{FeatureAlg} in Appendex \ref{ChapterstepExample}  is used to determine these features.
 	\item Rubine Feature Set (FS2) [12 features] .  Compute Rubine features as in \cite{gestureexample12} 
	\item Statistical Features (FS3) compute the moments as in \cite{zernike61} for N=10 there are 32 features.
	\item Global shape properties set (FS4). Firstly compute the convex hull of the stroke points then compute the rest of the features [12 features]. 
	\item Introduce the computed feature vector to the classifier and check the label of the detected symbol. 
\end{enumerate}

\begin{landscape}

More details are in Table \ref{tab:StepsRegDetails} in Appendex \ref{ChapterstepExample}

\begin{scriptsize}

 \begin{longtable}{|p{2cm}|p{2cm}|p{5cm}|p{2cm}|p{10cm}|}
\caption{Detailed Output of System in Each Step of Recognition}
\label{tab:StepsReg} \\

\hline 
\multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
\multicolumn{1}{p{2cm}|}{\textbf{Description}} &
\multicolumn{1}{p{5cm}|}{\textbf{Inputs}} &
\multicolumn{1}{p{2cm}|}{\textbf{Details}} &
\multicolumn{1}{p{10cm}|}{\textbf{Output}} 
\\ \hline 
\endfirsthead

\multicolumn{3}{c}%
{{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\  
\hline
\multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
\multicolumn{1}{p{2cm}|}{\textbf{Description}} &
 \multicolumn{1}{p{5cm}|}{\textbf{Inputs}} &
 \multicolumn{1}{p{2cm}|}{\textbf{Details}} &
 \multicolumn{1}{p{10cm}|}{\textbf{Output}} 
\\ \hline 
\endhead

 Feature Extraction of FS1 & Preprocessing of Structural Featues  &  \includegraphics[scale=0.5]{afterDefense/FinalSegmentation.jpg.eps}
 &  (see First Algorithm in Section \ref{sec:steprec} )  &    Line 1    is\textbf{ parallel} to Line from  L3

 Line 1  \textbf{ intersects }Line 2
  Line 2  \textbf{ intersects }Line 3
   Line 3  \textbf{ intersects }Line 4
    Line 4  \textbf{ intersects }Line 5

\\ \hline

 FS1 & Structural and geometrical Features & Segment List $L_s$  &  & 
 The final Feature vector of the FS1==>
ature Number of primitives 3.0   ,    Feature Number of segments 9.0   ,    Feature Line count 6.0   ,    Feature Curves Count 3.0   ,    Feature Ellipse Count 1.0   ,    Feature Intersection T type -1.0   ,    Feature Intersection L type 4.0   ,    Feature Intersection X type -1.0   ,    Feature Parallel Count 1.0   ,    Feature PerPendicular Count -1.0   ,    Feature Intersection Lines Count 4.0   ,    Feature Min Radius 1.7976931348623157E308   ,    Feature Max Radius 231.70851748053468

 \\ \hline
FS2  & Rubine Features & Segment List $L_s$  &   see \cite{gestureexample12}   & 
\begin{scriptsize}
 Feature Rubine 0 -1.0   ,    Feature Rubine 1 -1.0   ,    Feature Rubine 2 477.6107201476952   ,    Feature Rubine 3 0.6368010415482491   ,    Feature Rubine 4 169.49926253526885   ,    Feature Rubine 5 0.8908593331996381   ,    Feature Rubine 6 0.454279262624981   ,    Feature Rubine 7 1179.0172885029808   ,    Feature Rubine 8 4.712388980384688   ,    Feature Rubine 9 70.80157037633609   ,    Feature Rubine 10 55.4044569978846   ,    Feature Rubine 11 9.140625   ,    Feature Rubine 12 41783.0   ,   
\end{scriptsize}
 \\ \hline

FS3 & Momments Computation   &  Segment List $L_s$  &  see \cite{zernike61}  &

\begin{scriptsize}
   Feature Zernike moments 0 158.87520185460517   ,    Feature Zernike moments 0 7.007557168632727   ,    Feature Zernike moments 1 34.26844061831789   ,    Feature Zernike moments 1 6.658086415801404   ,....... ...... .... ....... ....... ....... ..... ..... ....... ....... ....... ..... .....
\end{scriptsize}

  \\ \hline
FS4   &  &Segment List $L_s$   &  &  \begin{scriptsize} 
 The convex hull    [ ]
 Area of convex hull =  -34160.5   preimiter of hull 656.1576413855265

 --------------------------------------------------------------------------------

Final Feature vector for the FS4: 
  Feature Centroid time 1.2504525968640884E12  , Feature Centroid time difference 28319.088339222613   ,    Feature area of convexhull 34160.5   ,    Feature area of convexhull/area of symbol 1.2365344240932454   ,    Feature N.Points ConvexHull/ N. points symbol 0.15302491103202848   ,    Feature Convext Perimeter/symbol perimeter 0.556529278903672   ,    ....... ....... ....... ..... ..... ....... ....... ....... ..... .....
  \end{scriptsize}
    \\ \hline
 Feature Vector & Appends the selected Feature FS1, FS2, FS3, FS4 & Previous Steps  & &

  [FS1,FS2,FS3,FS4]  \\ \hline
   Classification  & SVM classifier decision & Feature Vector  &  & 
   Symbol = Clock.

 \\ \hline

### 2009-09-06 draft/chapterSurvey.tex
after the changes in examples

\subsection{The Algorithm}
\label{sec:ParticleSwarmAlgorithm}

This section will discribe each step of the PSO Algorithm. Intially, the swarm partilces are created and each of which is a binary vector crossponding to candidate solution. Additionally, the velocities of each particle is initialized at random by drawing uniformly real values from the range [0.0 to 1.0]. Then the evolutionary process of particle movement is iterated until maximum number of eteration or a fittness error achieved. At each iteration,  the fitness value of each partilce is evaluated, and the particles individual local best and global best is updated. The particle position is adjusted as follow; the velocities is computed using Equation \ref{eq:Swarm}. Equation \ref{eq:vmax} is used to scale the velocity to values between 0 and 1 which will serve asa probability that the corresponding bit takes the value of 1. The particles bit vectors are updated by Equation \ref{eq:descrite} according to their velocity which measure the trajectory to the target solution based on local experience ($lbest_{i}$) and global intelligence ($gbest_{i}$). Before the end of the iteratin the particle may move to a refined location based on the refine procedure used.

\begin{verbatim}
Initialize 
  1. Generate N particles with M dimensions (P_i...P_N) 
  2. Generate velocities for each particle for M dimension randomly. 
  3. Repeat till Maximum number of iteration OR ErrorThreshold  
         For each particle P_i do the following 
         3.a   Compute fitness  (fitness)
         3.b   Determine Lbest  
               if (lbest_fitness<fitness)
                  3.b.1   lbest=P_i
         3.c      Determine Gbest.
               if (lGest_fitness<fitness)
                  3.c.1   lGest=P_i
         3.d   Compute the Velocity for each dimension using eq.3.2 
         3.e   Bound Velocity using eq.3.4
         3.f   Update particle values using eq. 3.3
         3.g   Refine the location of the particle
         end .
     end 
4. return Gbest
\end{verbatim}

>>>>

\subsection{The DPSO Algorithm}
\label{sec:ParticleSwarmAlgorithm}

This section will discribe each step of the PSO Algorithm. Figure\ref{fig:BlockPSO} illustrates the DPSO algorithm in Algorithm \ref{DPSOalg} steps in details. Initially, the swarm particles are created and each of which is a binary vector crossponding to candidate solution. Additionally, the velocities of each particle is initialized at random by drawing uniformly real values from the range [0.0 to 1.0]. Then the evolutionary process of particle movement is iterated until maximum number of iteration or a fitness error achieved. At each iteration,  the fitness value of each particle is evaluated, and the particles individual local best and global best is updated. The particle position is adjusted as follow; the velocities is computed using Equation \ref{eq:Swarm}. Equation \ref{eq:vmax} is used to scale the velocity to values between 0 and 1 which will serve as a probability that the corresponding bit takes the value of 1. The particles bit vectors are updated by Equation \ref{eq:descrite} according to their velocity which measure the trajectory to the target solution based on local experience ($lbest_{i}$) and global intelligence ($gbest_{i}$). Before the end of the iteration the particle may move to a refined location based on the refine procedure used. 
\begin{algorithm}
\caption{ Discrete Particle Swarm Algorithm }
\label{DPSOalg}
\begin{verbatim}
Initialize 
  1. Generate N particles with M dimensions (P_i...P_N) 
  2. Generate velocities for each particle for M dimension randomly. 
  3. Repeat till Maximum number of iteration OR ErrorThreshold  
         For each particle P_i do the following 
         3.a   Compute fitness  (fitness)
         3.b   Determine Lbest  
               if (lbest_fitness<fitness)
                  3.b.1   lbest=P_i
         3.c      Determine Gbest.
               if (lGest_fitness<fitness)
                  3.c.1   lGest=P_i
         3.d   Compute the Velocity for each dimension using eq.3.2 
         3.e   Bound Velocity using eq.3.4
         3.f   Update particle values using eq. 3.3
         3.g   Refine the location of the particle
         end .
     end 
4. return Gbest
\end{verbatim}
  \end{algorithm}

\begin{figure}
	\centering
		\includegraphics[scale=0.7]{afterDefense/BlockPSO.eps}
	\caption{Flow Chart of PSO Algorithm }
	\label{fig:BlockPSO}
\end{figure}
